Weak interspecific interactions in a sagebrush steppe: evidence from observations, models, and experiments

### 2017-08-29 removals/manuscript/ResponseToReviews.tex

\response{I think the real question about parameter uncertainty is this: if you sample a parameter vector at random from the posterior distribution, and re-make Fig. 5 using that parameter vector
in all years, would the results always look like Fig. 5, or are there other possible outcomes within the set of plausible parameter values? This would be a lot of work, and I don't see any good way to
summarize it graphically. }

My last comment, and really the only negative thing I can say about this paper is that it comes across as rather confirmatory in nature. It is absolutely solid science and highly interesting and important but the writing does not punch you with novelty. I think some small changes in wording could highlight that this is quite novel and hasn't been assessed in any other systems.

\response{Seems like the novelty is in the approach--experimental validation of a model based on observational data. We could also point out that while 
removal experiments have a long tradition, they are rarely coupled with a population model and generate only short term results like changes in cover or 
fitness components. Other thoughts?}

\response{SPE: here are a few more ideas, or maybe expansions on Peter's ideas and those in the paper. According to \texttt{http://onlinelibrary.wiley.com/doi/10.1111/cobi.12049/abstract}, 
population models fitted to observational data don't do well at out-of-sample prediction, because
conditions change and consequently out-of-sample environments are different from in-sample environments. Here we have a rare, perhaps unique, opportunity to (a) test the out-of-sample 
predictive power of models fitted to observational data when effects of changing environmental conditions are accounted for (in our case, by using the control plots to estimate random year effects), 
and (b) test the specific finding of Adler et al. (2010) that interspecific competition is very weak, relative to intraspecific, in this system. Given the conflicting results 
about the relative importance of within- and between-species competition, as reviewed in the manuscript, 
it is very informative to test the Adler et al. (2010) specific conclusions and thereby test the reliability of conclusions from models fitted to observational
data. Maybe it would also be worth talking about the fact that experimental comparisons of inter- vs intra-specific competition (such as Pantastico-Caldas, M. and D. L. Venable. 1993. 
Competition in two species of desert annuals along a topographic gradient. Ecology 74: 2192-2203) often (typically?) reach the opposite conclusion from observational studies (i.e, experiments
suggest that within- and between-species competition are comparable in strength). Are experimental tests too artificial, not capturing the processes that actually make inter$\gg$intra, or are 
models based on observational data too unreliable? The answer is important for our understanding of how plant communities operate, and for how we should study plant communities in the future.
The more we can say about our results having general methodological conclusions for future work, the more it will do for our ``general significance'' score.}

@@

\response{Not sure what to say about when models will under/over estimate interaction effects. Thoughts?}

One concern in the design is based on the relatively low covers, especially of the grasses. Based on
Figure 1, basal grass cover per species is very low even before removal—1-2\% per species. Finding
significant effects of such low cover on shrubs would indeed be surprising. Given this low expectation,
this does not seem a very strong test of whether model predictions are valid or not for the shrub. Does
the model even predict any effect of grass removals at these levels? \response{Yes, I can look these up}. Should the grass removal
treatment even be included in the study? The parameters estimated for A. tripartita seem to be very
uncertain in any case (pp 16-17), further suggesting that the results for this species do not add the story.
The test of the effects of shrubs on grasses is stronger (mean initial shrub cover was ca. 30\% before
removal) and indeed two out of three tests found significant shrub removal effects on growth beyond
the regressions on neighborhood crowding, as well as on population growth and the removal model
provides better predictions from the IBM for P. spicata. However, the authors argued this result may
not reflect stronger interspecific competition than estimated by the model because the release was no
greater for plants originally under a canopy than outside a sagebrush canopy. However, that argument
only makes sense either if competition is only for light or if the canopy edge reflects the edge of the
rooting zone as well. Given the habitat, I doubt the former is true and the authors also suggest it is
unlikely (line 442). For the latter, is anything known about the relationship of root distribution to
canopy cover in A. tripartita? My guess is that roots extend well beyond the canopy edge in these
habitats, leaving underestimated competitive interactions as a likely explanation for the significant
removal effect.

\response{Low cover should be irrelevant as long as the populations are near equilibrium. (SPE: why? Is the logic that
the cover at equilibrium, whatever it might be, creates enough competitive pressure to keep each of the species
from increasing?) I can remind the reviewer the grasses are measured with basal cover (SPE: do you also have data, from the
contemporary quadrats, that would let you translate basal cover into canopy cover for the grasses? In the pictures I've seen,
the total canopy cover for the grasses is higher than the canopy cover of ARTR. In fact, I think it might be useful to add to the
SI a figure consisting of maybe 4 or 9 typical quadrats photographed from above. People reading the article seem to think (incorrectly) 
that the community is mostly bare ground with the occasional plant here and there, and they are (partially mis-)interpreting our Results in light
of that belief.)}

@@

Line 87: given that you earlier defined competitive release as a population response, should
“competitive release” at the end of this sentence be replaced by “individual response to neighbors”?
The second question more directly addresses the population consequences.

Line 216: “very similar to those used in our previous studies” is a bit confusing. Why not just use the
vital rate functions calculated in the first part of the study? If they are indeed the same, say so rather
than introducing similarity to other published studies again.

Line 234-235: Several confusing points here: a) The distinction between the community reaching
equilibrium (500 time steps) and each species reaching equilibrium cover (“additional 2000 time steps”)
is not clear. b) I assume the random year effects were used for all 2500 time steps? c) were the 3
scenarios applied AFTER the community reached equilibrium but before the 2000 steps, i.e., removal
from an equilibrium community? Or from the beginning of each set of runs?

\response{SPE [I just went ahead and made some changes in the manuscript]: a) This was bad phrasing on our part. 
The community equilibrium is not a constant level of cover for each species, as it would be in a deterministic
model, but a steady-state pattern of fluctuations over time. The additional 2000 time steps are needed to compute the average cover of each species
at the community steady-state. We have re-phrased and expanded these sentences to clarify this point. b) Yes, that's right. The paragraph
has been reorganized to clarify this point. c) The scenarios were applied from the very beginning; the paragraph has been re-organized to 
clarify this point.}
